### 2008-10-10 trunk/Papers/rmrs-08/introduction.tex
titlebox back to normal

Representing semantics as a logical form that supports automated
inference and model construction is vital for deeper language
engineering tasks, such as dialogue systems.  Logical forms
can be obtained from hand-crafted deep grammars
\cite{butt:etal:1999,copestake:flickinger:2000}, but this
tends to lack robustness: they fail to cover all
words and constructions and by design don't handle
ill-formed phrases.

There has thus been a trend recently towards
robust wide-coverage semantic construction (e.g., 
\cite{bos:etal:2004,zettlemoyer:collins:2007}).     
% COMMENT: this issue about optional arguments is really too cryptic.
% Basically, there is no way for assessing by *any* learned language
% processor whether a single word form that appears in
% a number of different syntactic valencies is an example of lexical
% sense ambiguity, or they are all examples of the same sense being
% used in syntactically different ways (i.e., we have optional
% arguments).  Only hand-crafted grammars can tell you this.  So, for
% example, Bos et al (2004) would get bet(x,y) for "I bet 5 pounds",
% and bet(x,y,z) for "I bet you 5 pounds" and bet(x,p) for "I bet it's
% going to rain": I.e., the vocab of the semantic formalism is
% constructed automatically from NL word forms, and the predicate
% symbols in that automatically constructed vocab vary in (a) valency,
% and (b) the semantic dependencies represented by a given argument
% position (e.g., the second argument of "bet" in "I bet 5 pounds" is
% 5 pounds, and for "I bet it's raining" it's "it's raining".
% 
But there are certain semantic phenomena that these robust
approaches don't capture reliably, including quantifier scope,
optional arguments, and long-distance dependencies (for instance,
\newcite{clark:etal:2004} report that the parser used by
\newcite{bos:etal:2004} yields 63\% accuracy on object
extraction; e.g., {\em the man that I met\dots}).  Forcing a robust
parser to make a decision about these phenomena can therefore be
error-prone.  Depending on the application, it may be preferable to
give the parser the option to leave a semantic decision open when it's not
sufficiently informed---i.e., to compute a partial semantic
representation and to complete it later, using information
extraneous to the parser.

In this paper, we focus on an approach to representing partial
semantic information that supports this strategy: Robust Minimal
Recursion Semantics (\rmrs, \cite{copestake:2007a,copestake:2007b}).
\rmrs\ is designed to support underspecification of lexical information,
scope, and predicate-argument structure.  It is an emerging standard
for representing partial semantic information, and has been applied in
several implemented systems.  For instance, \newcite{copestake:2003}
and \newcite{frank:2004} use it to specify semantic components to
shallow parsers ranging in depth from {\sc pos} taggers to chunk
parsers and intermediate parsers such as {\sc rasp}
\cite{briscoe:etal:2006}.  {\sc mrs} analyses
\cite{copestake:etal:2005} derived from deep grammars, such as the
English Resource Grammar ({\sc erg}, \cite{copestake:flickinger:2000})
are special cases of {\sc rmrs}.

The key contribution we make is to cast {\sc rmrs}, for the first
time, as a logic with a well-defined model theory.  Previously, no
such model theory existed, and so \rmrs\ had to be used in a somewhat
ad-hoc manner which left open what exactly any given \rmrs\
representation actually \emph{means}.  This has hindered quite
practical progress, both in terms of understanding the relationship of
\rmrs\ to other frameworks such as \mrs\ and predicate logic and in
terms of the development of efficient algorithms.  As one application
of the formal concepts that arise from our definitions, we use
entailment to propose a
novel way of characterising consistency of \rmrs\ analyses across
different parsers.

>>>>

Representing semantics as a logical form that supports automated
inference and model construction is vital for deeper language
engineering tasks, such as dialogue systems.  Logical forms
can be obtained from hand-crafted deep grammars
\cite{butt:etal:1999,copestake:flickinger:2000} but 
this lacks robustness: not all
words and constructions are covered and by design
ill-formed phrases fail to parse. 
There has thus been a trend recently towards
robust wide-coverage semantic construction (e.g., 
\cite{bos:etal:2004,zettlemoyer:collins:2007}).     
% COMMENT: this issue about optional arguments is really too cryptic.
% Basically, there is no way for assessing by *any* learned language
% processor whether a single word form that appears in
% a number of different syntactic valencies is an example of lexical
% sense ambiguity, or they are all examples of the same sense being
% used in syntactically different ways (i.e., we have optional
% arguments).  Only hand-crafted grammars can tell you this.  So, for
% example, Bos et al (2004) would get bet(x,y) for "I bet 5 pounds",
% and bet(x,y,z) for "I bet you 5 pounds" and bet(x,p) for "I bet it's
% going to rain": I.e., the vocab of the semantic formalism is
% constructed automatically from NL word forms, and the predicate
% symbols in that automatically constructed vocab vary in (a) valency,
% and (b) the semantic dependencies represented by a given argument
% position (e.g., the second argument of "bet" in "I bet 5 pounds" is
% 5 pounds, and for "I bet it's raining" it's "it's raining".
% 
But there are certain semantic phenomena that these robust
approaches don't capture reliably, including quantifier scope,
optional arguments, and long-distance dependencies (for instance,
\newcite{clark:etal:2004} report that the parser used by
\newcite{bos:etal:2004} yields 63\% accuracy on object
extraction; e.g., {\em the man that I met\dots}).  Forcing a robust
parser to make a decision about these phenomena can therefore be
error-prone.  Depending on the application, it may be preferable to
give the parser the option to leave a semantic decision open when it's not
sufficiently informed---i.e., to compute a partial semantic
representation and to complete it later, using information
extraneous to the parser.

In this paper, we focus on an approach to
semantic representation that supports this strategy: Robust Minimal
Recursion Semantics (\rmrs, \newcite{copestake:2007a}).
\rmrs\ is designed to support underspecification of lexical information,
scope, and predicate-argument structure.  It is an emerging standard
for representing partial semantics, and has been applied in
several implemented systems.  For instance, \newcite{copestake:2003}
and \newcite{frank:2004} use it to specify semantic components to
shallow parsers ranging in depth from {\sc pos} taggers to chunk
parsers and intermediate parsers such as {\sc rasp}
\cite{briscoe:etal:2006}.  {\sc mrs} analyses
\cite{copestake:etal:2005} derived from deep grammars, such as the
English Resource Grammar ({\sc erg}, \cite{copestake:flickinger:2000})
are special cases of {\sc rmrs}.

The key contribution we make is to cast {\sc rmrs}, for the first
time, as a logic with a well-defined model theory.  Previously, no
such model theory existed, and so \rmrs\ had to be used in a somewhat
ad-hoc manner that left open exactly what any given \rmrs\
representation actually \emph{means}.  This has hindered 
practical progress, both in terms of understanding the relationship of
\rmrs\ to other frameworks such as \mrs\ and predicate logic and in
terms of the development of efficient algorithms.  As one application
of our formalisation, we use
entailment to propose a
novel way of characterising consistency of \rmrs\ analyses across
different parsers.

### 2008-10-10 trunk/Papers/rmrs-08/motivation.tex
titlebox back to normal

Consider the
following (toy) sentence:

\begin{examples}
  \item Every fat cat chased some dog.
\end{examples}

It exhibits several kinds of ambiguity, including a quantifier scope
ambiguity and lexical
ambiguities---e.g., the nouns ``cat'' and ``dog'' have 8 and 7
WordNet senses respectively.  Simplifying slightly by ignoring tense
information, two of its readings are shown as logical forms below;
these can be represented as trees as shown in Fig.~\ref{fig:1}.
%  (but see Section~\ref{sec:extensions})

@@

Now imagine we are trying to extract semantic information from the
output of a part-of-speech tagger by using the word lemmas as lexical
predicate symbols.  Such a semantic representation is highly partial.
It will use predicate symbols such as
$\sempred{\_cat\_n}$, which might resolve to the predicate symbols
$\sem{\_cat\_n\_1}$ or $\sem{\_cat\_n\_2}$ in the complete semantic
representation.  (Notice the different fonts for the ambiguous and
unambiguous predicate symbols.)
But
most underspecification formalisms (e.g., {\sc mrs}
\cite{copestake:etal:2005} and {\sc clls} \cite{egg:etal:2001}) are
unable to represent semantic 
information that is as partial as what we get from a {\sc pos} tagger
because they cannot underspecify
predicate-argument structure.
\rmrs\ \cite{copestake:2007a} was
designed to address this problem.  In \rmrs, we can express the
information we get from the {\sc pos} tagger as follows:

@@

This \rmrs\ expresses only that certain predications must be made in
the semantic representation---but it doesn't say anything about
semantic scope, about most arguments of the
predicates (note that $\sempred{\_chase\_v}(e)$ doesn't say who chases
whom), or about the coindexation of variables ($\sempred{\_every\_q}$
binds the variable $x_1$, whereas $\sempred{\_cat\_n}$ speaks about
$x_3$), and it maintains the lexical ambiguities.  Technically, it
consists of six \emph{elementary predications} ({\sc ep}s), one for
each word lemma in the sentence; each of them is prefixed by a
\emph{label} and an \emph{anchor}, which are essentially variables
that refer to nodes in the trees in Fig.~\ref{fig:1}.  We can say that
the two trees \emph{satisfy} this \rmrs\ because it is possible to map
the labels and anchors into nodes in each tree and variable names like
$x_1$ and $x_3$ into variable names in the tree in such a way that the
predications of the nodes that labels and anchors are mapped to are 
consistent with those in the {\sc ep}s of the {\sc rmrs}.  For
instance, the first tree satisfies the \rmrs\ because $l_1$ and $a_1$
can map to the root of that tree and $x_1$ to $x$, and the root label
$\sem{\_every\_q\_1}$ is consistent with the {\sc ep} predicate
$\sempred{\_every\_q}$.

@@

There are of course many other trees (and thus, fully specific
semantic representations such as (\ref{ex:fat-cat-1})) that are
described equally well by the \rmrs\ (\ref{ex:cat-pos}); this is not
surprising, given that the semantic information we got from the {\sc
  pos} tagger is so incomplete.  If we have
information about subjects and objects from a chunk parser
like Cass \cite{abney:1996}, we can represent it in a more detailed
\rmrs:

@@

This introduces two new types of atoms.   $x_3=x_4$
means that $x_3$ and $x_4$ map to the same
variable in any fully specific logical form; e.g., both to the
variable $x$ in Fig.~\ref{fig:1}.   $\ARG_i(a,z)$ and
$\ARG_i(a,h)$ express that the $i$-th child (counting from 0) of the
node to which the anchor $a$ refers is the variable name that $z$
denotes (or the node that the {\em hole} $h$ denotes).  So unlike
earlier underspecification formalisms, \rmrs\ can specify the
predicate of an atom separately from its
arguments; this is necessary for supporting parsers where information
about lexical subcategorisation is absent. If we also allow atoms of
the form $\ARG_{\{2,3\}}(a,x)$ to express uncertainty as to whether
$x$ is the second or third child of the anchor $a$, then \rmrs\ can
even specify the arguments to a predicate while underspecifying their
position.  This can useful for specifying arguments to \_give\_v when a
parser doesn't handle unbounded dependencies and is faced with
{\em Which bone did you give the dog?} vs.\ {\em To which dog did you
  give the bone?}

@@

$\mathsf{RSTR}$ and $\mathsf{BODY}$ are conventional names
for the $\ARG_1$ and $\ARG_2$ of a quantifier predicate symbol.  Atoms
like $h_2 \qeq l_{42}$ (``qeq'') specify a certain kind of
``outscopes'' relationship between the hole and the label, and are
used here to underspecify the scope of the two quantifiers.

Notice that the labels of the {\sc ep}s for ``fat'' and ``cat'' are
stipulated to be equal in this \rmrs, whereas the anchors are not.  In
the tree, it is the anchors that are mapped to the nodes with the
labels $\sem{\_fat\_j\_1}$ and $\sem{\_cat\_n\_1}$; the label is
mapped to the conjunction node just above them.  In other words, the
role of the anchor in an {\sc ep} is to connect a predicate to its
arguments, while the role of the label is to connect the {\sc ep} to
the surrounding formula.  Representing conjunction with label sharing
stems from \mrs\ and provides compact representations.

Finally, (\ref{ex:cat-erg}) uses predicate symbols like
$\sempred{\_dog\_n\_1}$ that are meant to be more specific than the
symbols like $\sempred{\_dog\_n}$ which the earlier \rmrs s used.
This reflects the fact that the deep grammar will perform some lexical
disambiguation that the chunker and {\sc pos} tagger didn't.  The fact
that the former symbol should be more specific than the latter can be
represented using SPEC atoms like $\sempred{\_dog\_n\_1}\sqsubseteq
\sempred{\_dog\_n}$.
Note that even a deep grammar will not fully disambiguate to
semantic predicate symbols, such as WordNet senses, and so
$\sempred{\_dog\_n\_1}$ can still be consistent with multiple symbols
like $\sem{\_dog\_n\_1}$ and $\sem{\_dog\_n\_2}$ in the semantic
representation.  However, unlike the output of a
{\sc pos} tagger, an \rmrs\ symbol that's output by a deep grammar is
consistent 
with symbols that all have the {\em same} arity, because a deep grammar
fully determines lexical subcategorisation.

In summary, \rmrs\ allows us to represent the (partial) semantic information
that can be extracted from a wide range of {\sc nlp} tools in a uniform way.
This is useful for hybrid systems which fall back to shallower
analyses if deeper parsing is unsuccessful, or which try to match
deeply parsed queries against shallow parses of large text corpora;
and in fact, \rmrs\ is gaining popularity as a practical interchange
format for exactly these purposes \cite{copestake:2003,frank:2004}.
However, \rmrs\ is still relatively ad-hoc in that its
formal semantics is
not defined; we don't know, formally, what an \rmrs\ \emph{means} in
terms of semantic representations like (\ref{ex:fat-cat-1}) and
(\ref{ex:fat-cat-2}), and this hinders our ability to design efficient
algorithms for processing \rmrs. The purpose of this paper is to lay
the groundwork for fixing this problem.

>>>>

Consider the
following (toy) sentence:
\begin{examples}
\label{ex:cat}
  \item Every fat cat chased some dog.
\end{examples}
It exhibits several kinds of ambiguity, including a quantifier scope
ambiguity and lexical
ambiguities---e.g., the nouns ``cat'' and ``dog'' have 8 and 7
WordNet senses respectively.  Simplifying slightly by ignoring tense
information, two of its readings are shown as logical forms below;
these can be represented as trees as shown in Fig.~\ref{fig:1}.
%  (but see Section~\ref{sec:extensions})

@@

Now imagine trying to extract semantic information from the output of
a part-of-speech ({\sc pos}) tagger by using the word lemmas as
lexical predicate symbols.  Such a semantic representation is highly
partial.  It will use predicate symbols such as $\sempred{\_cat\_n}$,
which might resolve to the predicate symbols $\sem{\_cat\_n\_1}$ or
$\sem{\_cat\_n\_2}$ in the complete semantic representation.  (Notice
the different fonts for the ambiguous and unambiguous predicate
symbols.)  But most underspecification formalisms (e.g., {\sc mrs}
\cite{copestake:etal:2005} and {\sc clls} \cite{egg:etal:2001}) are
unable to represent semantic information that is as partial as what we
get from a {\sc pos} tagger because they cannot underspecify
predicate-argument structure.  \rmrs\ \cite{copestake:2007a} is 
designed to address this problem.  In \rmrs, the
information we get from the {\sc pos} tagger is as follows:

@@

This \rmrs\ expresses only that certain predications are present in
the semantic representation---it doesn't say anything about
semantic scope, about most arguments of the
predicates (e.g., $\sempred{\_chase\_v}(e)$ doesn't say who chases
whom), or about the coindexation of variables ($\sempred{\_every\_q}$
binds the variable $x_1$, whereas $\sempred{\_cat\_n}$ speaks about
$x_3$), and it maintains the lexical ambiguities.  Technically, it
consists of six \emph{elementary predications} ({\sc ep}s), one for
each word lemma in the sentence; each of them is prefixed by a
\emph{label} and an \emph{anchor}, which are essentially variables
that refer to nodes in the trees in Fig.~\ref{fig:1}.  We can say that
the two trees \emph{satisfy} this \rmrs\ because it is possible to map
the labels and anchors in (\ref{ex:cat-pos}) into nodes in each tree
and variable names like 
$x_1$ and $x_3$ into variable names in the tree in such a way that the
predications of the nodes that labels and anchors denote are 
consistent with those in the {\sc ep}s of (\ref{ex:cat-pos}).  For
$l_1$ and $a_1$
can map to the root of the first tree in Fig.~\ref{fig:1} and $x_1$ to
$x$, and the root label 
$\sem{\_every\_q\_1}$ is consistent with the {\sc ep} predicate
$\sempred{\_every\_q}$.

@@

There are of course many other trees (and thus, fully specific
semantic representations such as (\ref{ex:fat-cat-1})) that are
described equally well by the \rmrs\ (\ref{ex:cat-pos}); this is not
surprising, given that the semantic output from the {\sc
  pos} tagger is so incomplete.  If we have
information about subjects and objects from a chunk parser
like Cass \cite{abney:1996}, we can represent it in a more detailed
\rmrs:

@@

This introduces two new types of atoms.   $x_3=x_4$
means that $x_3$ and $x_4$ map to the same
variable in any fully specific logical form; e.g., both to the
variable $x$ in Fig.~\ref{fig:1}.   $\ARG_i(a,z)$ (and
$\ARG_i(a,h)$) express that the $i$-th child (counting from 0) of the
node to which the anchor $a$ refers is the variable name that $z$
denotes (or the node that the {\em hole} $h$ denotes).  So unlike
earlier underspecification formalisms, \rmrs\ can specify the
predicate of an atom separately from its
arguments; this is necessary for supporting parsers where information
about lexical subcategorisation is absent. If we also allow atoms of
the form $\ARG_{\{2,3\}}(a,x)$ to express uncertainty as to whether
$x$ is the second or third child of the anchor $a$, then \rmrs\ can
even specify the arguments to a predicate while underspecifying their
position.  This can useful for specifying arguments to \_give\_v when a
parser doesn't handle unbounded dependencies and is faced with
{\em Which bone did you give the dog?} vs.\ {\em To which dog did you
  give the bone?}

@@

$\mathsf{RSTR}$ and $\mathsf{BODY}$ are conventional names
for the $\ARG_1$ and $\ARG_2$ of a quantifier predicate symbol.  Atoms
like $h_2 \qeq l_{42}$ (``qeq'') specify a certain kind of
``outscopes'' relationship between the hole and the label, and are
used here to underspecify the scope of the two quantifiers.
Notice that the labels of the {\sc ep}s for ``fat'' and ``cat'' are
stipulated to be equal in (\ref{ex:cat-erg}), whereas the anchors are not.  In
the tree, it is the anchors that are mapped to the nodes with the
labels $\sem{\_fat\_j\_1}$ and $\sem{\_cat\_n\_1}$; the label is
mapped to the conjunction node just above them.  In other words, the
role of the anchor in an {\sc ep} is to connect a predicate to its
arguments, while the role of the label is to connect the {\sc ep} to
the surrounding formula.  Representing conjunction with label sharing
stems from \mrs\ and provides compact representations.

Finally, (\ref{ex:cat-erg}) uses predicate symbols like
$\sempred{\_dog\_n\_1}$ that are meant to be more specific than 
symbols like $\sempred{\_dog\_n}$ which the earlier \rmrs s used.
This reflects the fact that the deep grammar performs some lexical
disambiguation that the chunker and {\sc pos} tagger don't.  The fact
that the former symbol should be more specific than the latter can be
represented using SPEC atoms like $\sempred{\_dog\_n\_1}\sqsubseteq
\sempred{\_dog\_n}$.
Note that even a deep grammar will not fully disambiguate to
semantic predicate symbols, such as WordNet senses, and so
$\sempred{\_dog\_n\_1}$ can still be consistent with multiple symbols
like $\sem{\_dog\_n\_1}$ and $\sem{\_dog\_n\_2}$ in the semantic
representation.  However, unlike the output of a
{\sc pos} tagger, an \rmrs\ symbol that's output by a deep grammar is
consistent 
with symbols that all have the {\em same} arity, because a deep grammar
fully determines lexical subcategorisation.

In summary, \rmrs\ allows us to represent in a uniform way
the (partial) semantics
that can be extracted from a wide range of {\sc nlp} tools.
This is useful for hybrid systems which exploit shallower
analyses when deeper parsing fails, or which try to match
deeply parsed queries against shallow parses of large corpora;
and in fact, \rmrs\ is gaining popularity as a practical interchange
format for exactly these purposes \cite{copestake:2003}.
However, \rmrs\ is still relatively ad-hoc in that its
formal semantics is
not defined; we don't know, formally, what an \rmrs\ \emph{means} in
terms of semantic representations like (\ref{ex:fat-cat-1}) and
(\ref{ex:fat-cat-2}), and this hinders our ability to design efficient
algorithms for processing \rmrs. The purpose of this paper is to lay
the groundwork for fixing this problem.

### 2008-10-10 trunk/Papers/rmrs-08/rmrs.tex
titlebox back to normal

We define the syntax of \rmrs\ in the style of {\sc clls}
\cite{egg:etal:2001}.  We assume an infinite set of {\em node
  variables} $\nvar=\{X,Y,X_1,\ldots\}$, used as labels, anchors, and
holes; the distinction between these uses will come from their
position in the formula.  Furthermore, we assume an infinite set of
\emph{base variables} $\ovar$, consisting of individual variables
$\{x,x_1,y,\ldots\}$ and event variables $\{e_1,\ldots\}$.  Finally,
we assume a vocabulary of \emph{predicate symbols}
$\pred=\{P,Q,P_1,\ldots\}$.

We can now define formulas of \rmrs\ as follows.

@@

A node variable $X$ is called a \emph{label} iff $\varphi$ contains an
atom of the form $\ep{X}{Y}{P}$ or $Y \qeq X$; $X$ is called an
\emph{anchor} in the \rmrs\ $\varphi$ iff $\varphi$ contains an atom
of the form $\ep{Y}{X}{P}$ or $\ARG_S(X,i)$; and $X$ is called a
\emph{hole} iff $\varphi$ contains an atom of the form $\ARG_S(Y,X)$
or $X \qeq Y$.
\end{definition}

@@

Intuitively, we expect that an \rmrs\ in solved form is fully
specified with respect to the predicate-argument structure, all
variable equalities and inequalities and scope ambiguities have been
resolved, and only lexical sense ambiguities remain.  This is made
precise below.

@@

Because solved forms are so restricted, we can `read off' at least one
model from each solved form.

\begin{prop} \label{prop:solved-forms-are-satisfiable}
  Every {\sc rmrs} in solved form is satisfiable.
\end{prop}
\begin{proof}[Proof (sketch; see also
  \cite{Duchier00dominanceconstraints})] 
  For each {\sc ep}, we choose to label the anchor with the
  constructor $p$ of sufficiently high arity whose existence we
  assumed; we determine the edges between an anchor and its children
  from the uniquely determined $\ARG$ atoms; plugging labels into
  holes is straightforward because no label is dominated by more than
  one label; and then we can fill the spaces between the labels and
  anchors with conjunctions.
\end{proof}

@@

Another feature of the existing \rmrs\ literature is that each term of
an \rmrs\ is equipped with a \emph{sort}.  For instance, individual
variables are typically assigned the sort $x$, event variables the
sort $e$, and holes the sort $h$; these are arranged together with
their subsorts in a sort hierarchy ${\cal S}$.  While we have defined
\rmrs\ without sorts to simplify the presentation, it is
straightforward to add them.  For this, one assumes that the signature
$\Sigma$ is sorted, i.e.\ assigns a sort $s_1\times\ldots
s_n\rightarrow s$ to each constructor, where $n$ is the constructor's
arity (possibly zero) and $s,s_1,\ldots,s_n\in {\cal S}$ are atomic
sorts.  We restrict the models of \rmrs\ to trees that are
\emph{well-sorted} in the usual sense, i.e.\ those in which we can
infer a sort for each subtree, and require that the variable
assignment functions likewise respect the sorts.  If we then modify
Def.~\ref{defn:solved-forms} such that the constructor $p$ of
sufficiently high arity is also consistent with the sorts of the known
arguments---i.e., if $p$ has sort $s_1 \times \ldots \times s_n
\rightarrow s$ and the \rmrs\ contains an atom $\ARG_{\{k\}}(Y,i)$ and
$i$ is of sort $s'$, then $s'$ is a subsort of $s_k$---all the above
propositions about solved forms remain true.

>>>>

We define \rmrs\ syntax in the style of {\sc clls}
\cite{egg:etal:2001}.  We assume an infinite set of {\em node
  variables} $\nvar=\{X,Y,X_1,\ldots\}$, used as labels, anchors, and
holes; the distinction between these will come from their
position in the formulas.  We also assume an infinite set of
\emph{base variables} $\ovar$, consisting of individual variables
$\{x,x_1,y,\ldots\}$ and event variables $\{e_1,\ldots\}$, and a
vocabulary of \emph{predicate symbols}
$\pred=\{P,Q,P_1,\ldots\}$.
\rmrs\ formulas are defined as follows.

@@

A node variable $X$ is called a \emph{label} iff $\varphi$ contains an
atom of the form $\ep{X}{Y}{P}$ or $Y \qeq X$; it is an 
\emph{anchor} iff $\varphi$ contains an atom
of the form $\ep{Y}{X}{P}$ or $\ARG_S(X,i)$; and it is a
\emph{hole} iff $\varphi$ contains an atom of the form $\ARG_S(Y,X)$
or $X \qeq Y$.
\end{definition}

@@

Intuitively, an \rmrs\ in solved form is fully
specified with respect to the predicate-argument structure, all
variable equalities and inequalities and scope ambiguities have been
resolved, and only lexical sense ambiguities remain.  This is made
precise below.

@@

Because solved forms are so restricted, we can `read off' at least one
model from each solved form:
\begin{prop} \label{prop:solved-forms-are-satisfiable}
  Every {\sc rmrs} in solved form is satisfiable.
\end{prop}
\begin{proof}[Proof (sketch; see also
  \cite{Duchier00dominanceconstraints})] 
  For each {\sc ep}, we choose to label the anchor with the
  constructor $p$ of sufficiently high arity whose existence we
  assumed; we determine the edges between an anchor and its children
  from the uniquely determined $\ARG$ atoms; plugging labels into
  holes is straightforward because no label is dominated by more than
  one label; and  spaces between the labels and
  anchors are filled with conjunctions.
\end{proof}

@@

Another feature of the existing \rmrs\ literature is that each term of
an \rmrs\ is equipped with a \emph{sort}.  In particular, individual
variables $x$, event variables $e$ and holes $h$ are arranged together
with their subsorts (e.g., $e_{\mbox{past}}$) and supersorts (e.g.,
sort $i$ abstracts over $x$ and $e$) into a
sort hierarchy ${\cal S}$. 
For simplicity we defined \rmrs\ without sorts, but it is
straightforward to add them.  For this, one
assumes that the signature $\Sigma$ is sorted, i.e.\ assigns a sort
$s_1\times\ldots s_n\rightarrow s$ to each constructor, where $n$ is
the constructor's arity (possibly zero) and $s,s_1,\ldots,s_n\in {\cal
  S}$ are atomic sorts.  We restrict the models of \rmrs\ to trees
that are \emph{well-sorted} in the usual sense, i.e.\ those in which
we can infer a sort for each subtree, and require that the variable
assignment functions likewise respect the sorts.  If we then modify
Def.~\ref{defn:solved-forms} such that the constructor $p$ of
sufficiently high arity is also consistent with the sorts of the known
arguments---i.e., if $p$ has sort $s_1 \times \ldots \times s_n
\rightarrow s$ and the \rmrs\ contains an atom $\ARG_{\{k\}}(Y,i)$ and
$i$ is of sort $s'$, then $s'$ is a subsort of $s_k$---all the above
propositions about solved forms remain true.
